\chapter{Applications}
\label{chap:applications}


To illustrate the use of the graphical user interface, two different physical
applications are shown in the desktop and the web frontend, respectively. From
the physics point of view, the example in the desktop version focuses more on the
density of states and the visualization of spin contributions, while the dataset
in web frontend focuses on the band structure $E(k)$ and the visualization
of defect states in supercells.

\section{Web Frontend: $\textrm{MoSe}_2$ Crystal}

Figure \textbf{TODO} shows the visualization of a band structure calculation of a 3 dimensional Molybdenum diselenide ($\textrm{MoSe}_2$ bulk) crystal using the default settings of the GUI. Even with the default settings the band structure plots clearly indicate, that $MoSe_2$ is a semiconductor, since there are no states at the Fermi level. Because the minimum of the conduction band is located at an other $k$ as the maximum of the valence band, the plot shows that $\textrm{MoSe}_2$ has an indirect band gap. This indicates that for the transition with the smallest energy difference between valence and the conduction band, both, energy and momentum have to change.

In contrast to the 3 dimensional extended $MoSe_2$ crystal, a $MoSe_2$ monolayer (see Fig. \textbf{TODO}) has a direct band gap but is still a semiconductor. Furthermore, the $\textrm{MoSe}_2$ monolayer has a defect atom in every 9th unit cell and the DFT computation is therefore done in a $3 \times 3$ supercell to restore periodicity. This is the reason for the much greater number of states in the band structure plot. 

Since the Brillouin zone of the supercell is smaller than the Brillouin zone of the same crystal without the defect, the supercell Brillouin zone is unfolded to the same size as the Brillouin zone of the unperturbed lattice. To account for the fact that the defect is only present in every 9th cell and its relative importance for the spectrum is therefore degraded, an unfolding weight is introduced to visualize the relative importance of bands in the unfolded Brillouin zone. By default, the unfolding weight is used by our visualization tool, but it can gradually be turned off in order to highlight the impact of defect states. This is shown in figure \textbf{TODO}. To even better visualize the defect state, it would also be possible to select the atom group belonging the defect atoms only. In this example, the analysis with reduced band unfolding shows, that there are many more direct band gaps origination from the defect state.




\section{Desktop Frontend: $\textrm{Co}$ Crystal}


% bänder isonlieren schwierig....fleur update--> label der E eigenvlas

%$m^{*}$ and $v_{G}(\vec{k})$
\section{Derived physical Quantities: Differentiation}
The kind of datasets handled in the scope of the project did not lend themselves
readily to applications of automatic differentiation techniques. Nevertheless,
it is possible to derive meaningful physical quantities from the band structure
using numerical differentiation techniques.

The effective mass $m^{*}$ represents the mass, that an electron appears to have due to the inter-atomic forces in the crystal. At every $\vec{k}_0$, where $E(\vec{k})$ has a local extremum, $E(\vec{k})$ can be expanded in a Taylor series with a vanishing first order term $E(k) = E_0 + \frac{\partial E(\vec{k})}{\partial \vec{k}} \cdot (\vec{k}-\vec{k}_0)^2$. Comparing this to the dispersion relation of a free electron $E(\vec{k})_{free} = \frac{\hbar^2 \vec{k}^2}{2 m_e}$ motivates the general definition 

\begin{equation}
    m^{*}_{k_i,k_j} = \hbar^2  \left(\frac{\partial^2E(\vec{k})}{\partial k_i \partial k_j}\right)^{-1}
\end{equation}

Since $\frac{\partial^2E(k)}{\partial \vec{k}_i \partial \vec{k}_j}$ depends on the direction of the partial derivatives, $m^{*}\big|_{i,j}$ is a tensor. Because the band structure files only contain a discrete sampled path in the Brillouin zone, only the derivatives that correspond to the direction from one high-symmetry point to the next can be computed. We are only interested in the diagonal terms of $m^{*}\big|_{i,j}$.

%maybe something about the directional dependency --> m is tensor...


A second potentially interesting quantity is the group velocity $v_{G}(\vec{k})$ associated to each band. The group velocity $v_{G}(\vec{k})$ at the Fermi energy $E = E_F$ is called Fermi velocity.

\begin{equation}
    v_{G}(\vec{k})\big|_i = \frac{1}{\hbar}\frac{\partial E(\vec{k})}{\partial \vec{k_i}}
\end{equation}


\subsection{Differentiation}

Since the k-mesh in DFT calculations is potentially very sparse, low order finite difference schemes are not expected to work well. Alternatively, one way to exploit all data points efficiently is to use fast Fourier transform methods, which are equivalent to the derivation of a truncated Fourier series. This method is expected be well-suited for the problem since the graph of the band structure $(k, E(|k|))$ with $k \in \{-\Gamma,..., H,..., \Gamma\}$ is periodic, where $\Gamma$ and $H$ are arbitrary representatives of the high symmetry points. This periodicity in the reciprocal space is a direct consequence of the periodicity of the crystal.

In Fourier space, spatial derivatives transform into multiplications, which can easily be shown by partial integration.

\begin{equation}
    f^{(n)}(x) = \mathcal{F}^{-1}\left((ik)^n\mathcal{F}(f(x))\right)
\end{equation}
 
To test the FFT differentiation method, a band was selected that did not have intersections with other bands within the interval between two high symmetry points and a stationary point at the high symmetry points. Then a resolution study was done to investigate the impact of the number of points within the interval.

The comparison between the FFT and a first-order central difference approximation of the second derivative (FD) is shown in Fig. \textbf{TODO}, where different k-mesh resolutions are compared to a derivative that is almost fully converged. The comparison indicates that for small N, the error of the FFT method is significantly smaller than the error of the FD derivative. This is especially striking in the vicinity of the high symmetry points, where the error of the differentiation is required to be small in order to get good approximations for $m^{*}$, which is only meaningful close to the high symmetry points.

When using more points, the difference between both approximation schemes is not significant. It is interesting to note, that the FFT method does not work anymore in the limit of extremely many points. In this case the derivative is dominated by Gibbs oscillations. These are most likely caused by small discontinuities in $E(k)$, due to basis changes inside the DFT computation.

At the current state the question remains to be answered weather the computation of $m^{*}$ and $v_g$ is useful. The number of bands, which the computation can be applied to is extremely limited. Since the bands in the data files are not labeled according to their corresponding eigenfunction but sorted by value, there are discontinuities at each point, where two bands intersect. This problem might be solved in the future. 













 %   \item Effective mass, that an electon in a crystal appears to have compared to a free %electron (due to interactions in the solid)
%    $m^{*} = \hbar^2  \left(\frac{\partial^2E(k)}{\partial k^2}\right)^{-1}$
%    \
%    \item Group velocity:
%    $v_{G}(\vec{k}) = \frac{1}{\hbar}\frac{\partial E(\vec{k})}{\partial \vec{k}}$%
%
%    
%    \item Problem: sparse k-Point mesh, but periodic band structure
%    \item Idea: Using FFT to compute accurate derivates:
%    \newline $\Leftrightarrow$ Differentiate a finite Fourier series
%    \newline $f^{(n)}(x) = \mathcal{F}^{-1}\left((ik)^n\mathcal{F}(f(x))\right)$


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../report"
%%% End:
