\clearpage

%%%%%%%%%%%%%%%% 
%%% Abstract %%%
%%%%%%%%%%%%%%%% 

\thispagestyle{empty}

\vspace*{\fill}
\noindent\rule[2pt]{\textwidth}{0.5pt}\\
{\textbf{Abstract ---}}
%% JW ======================================== 
The electronic band structure and the density of states of a material can be used to study many of it's electrical, magnetic and optical properties. Band structures of real-world materials are quite complex and not accessible with analytic methods. One way to obtain them is numerical simulation using density functional theory (DFT) on high-performance computers.
%and high-performance computing. 
Still, the raw numerical data output by itself is not easy to interpret.
Interactive visualization as a post-processing step can aid to explore that
space and find the important features. In this work we develop a freely
available post-processing pipeline for band structure calculations of Fleur, a
realization of DFT based on the all-electron full potential linearized augmented
planewave method. It features a generalized Python interface for processing and
visualizing data in the hierarchical data format HDF5. Two graphical user
interfaces (GUIs) built on that interface are introduced. Interactive controls
let the user visually distinguish the effects of different atom groups,
orbitals and defect states in supercells on the band structure. Simulations of
$\textrm{MoSe}_2$ and $\textrm{Co}$ materials are examined using those GUIs,
helping to deduce some of their physical properties.

% %% PK ========================================
% In this project, the raw data of the DFT simulations from Fleur are read,
% pre-processed and stored in suitable variables, then extracted various features
% from the data from the DFT simulations such as Fermi velocity, Effective mass.
% Then this data is visualized using suitable color pots for a physicist to
% understand the plots, material characteristics using the plots. This API is
% converted into a frontend GUI for easier use in future. The overall project
% helps physicists to solve their problem of reading the DFT simulations data.

{\textbf{Keywords}}
%% JW ========================================
band structure, density of states, visualization, DFT simulation, Fleur, HDF5
\\

% %% PK ========================================
% Fleur, DFT Simulations, visualization, Front end GUI, Feature extraction.
% \\

\noindent\rule[2pt]{\textwidth}{0.5pt}
\begin{center}
    AICES\\
    Schinkelstr. 2\\
    Rogowski Building\\
    4th Floor\\
    52062 Aachen    
\end{center}
\vspace*{\fill}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "report"
%%% End:
